---
title: "EDA"
author: "Shane Hauck"
date: "2024-03-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages
```{r}
library(tidyverse)
library(deldir)

```

# Load StatsBomb Package
```{r}
library(StatsBombR)

```

# Load Data
```{r}
# Premier League 2015/16 Event Level Data
# Comp <- FreeCompetitions()
# Matches <- FreeMatches(Comp)
# Matches <- Matches %>% filter(competition.competition_id == 2 & season.season_id == 27)
# Events <- free_allevents(Matches)
# Events <- allclean(Events)

```

```{r}
# World Cup 2022 360 Freeze Frame Data
Comp <- FreeCompetitions()
Matches <- FreeMatches(Comp)
Matches <- Matches %>% filter(competition.competition_id == 43 & season.season_id == 106)
data360 <- free_allevents_360(MatchesDF = Matches, Parallel = T)

events <- free_allevents(MatchesDF = Matches, Parallel = T)
events <- allclean(events)
events <- get.opposingteam(events)

data360 <- data360 %>% rename(id = event_uuid)
events <- events %>% left_join(data360, by = c("id" = "id"))
events <- events %>%
  rename(match_id = match_id.x) %>%
  select(-match_id.y)
```



# Dataset with only relevant columns for research
```{r}
interestCols <- c(
  "id", "index", "match_id", 
  "period", "timestamp", "minute", "second", "possession",
  "duration", "TimeInPoss", "milliseconds", "ElapsedTime", "StartOfPossession", "TimeToPossEnd",
  "related_events",  
  "possession_team.id", "possession_team.name", "tactics.formation",
  "player.id", "player.name", "position.id", "position.name", 
   "type.id", "type.name", "play_pattern.id", "play_pattern.name", "location.x", "location.y", 
  "pass.length", "pass.angle", "pass.end_location.x", "pass.end_location.y", 
  "pass.outcome.id", "pass.outcome.name", "pass.recipient.id", "pass.recipient.name",
  "pass.height.id", "pass.height.name", "pass.type.id", "pass.type.name",
  "pass.body_part.id", "pass.body_part.name", "pass.technique.id", "pass.technique.name",
  "pass.switch", "pass.aerial_won", "pass.assisted_shot_id", "pass.shot_assist",
  "pass.cross", "pass.cut_back", "pass.through_ball", "pass.deflected", 
  "carry.end_location.x", "carry.end_location.y", "under_pressure",
  "ball_receipt.outcome.id", "ball_receipt.outcome.name", 
  "freeze_frame", "visible_area"
  
)
```

# Select for relevant columns
```{r}
events_rel <- events %>% select(all_of(interestCols))

```


# Sum duration of related events
```{r}
events_time <- events_rel %>% 
  filter(type.name %in% c("Carry", "Pass")) %>%
  group_by(match_id) %>%
  arrange(match_id, index) %>%
  mutate(index = row_number()) %>%
  select(id, index, related_events, duration, type.name, player.name) %>%
  # When player.name is equal for sequential index, sum the duration
  mutate(change = player.name != lag(player.name, default = player.name[1])) %>%
  mutate(group_id = cumsum(change)) %>%
  group_by(match_id, group_id) %>%
  reframe(id, totalTimeInPoss = sum(duration, na.rm = TRUE)) %>%
  ungroup()
  

```

# Merge totalTimeInPoss with events_rel
```{r}
events_rel <- events_rel %>% left_join(events_time, by = c("id", "match_id"))

```

# Get ball receipt information
```{r}
events_rel <- events_rel %>% filter(!is.na(group_id)) %>%
  group_by(match_id, group_id) %>%
  mutate(recipient.location.x = first(location.x),
         recipient.location.y = first(location.y)) %>% 
  mutate(carried = ifelse(n() > 1, TRUE, NA)) %>%
  ungroup()
```

# Keep only the passes
```{r}
events_rel <- events_rel %>% filter(type.name == "Pass")

```

# Use an xgboost to create expected pass completeion probability
```{r}
events_rel <- events_rel %>% 
  mutate(pass.completed = ifelse(!is.na(pass.outcome.id), 1, 0)) 

library(xgboost)

vars_to_remove <- c("index", "match_id", "period", "timestamp", "minute", "second", "TimeInPoss",
                    "possession", "milliseconds", "ElapsedTime", "StartOfPossession", "TimeToPossEnd",
                    "related_events", "possession_team.id", "possession_team.name", "tactics.formation",
                    "player.id", "player.name",  "position.name", "type.id", "type.name", "play_pattern.name",
                    "pass.outcome.id", "pass.outcome.name", "pass.recipient.id", "pass.recipient.name",
                    "pass.height.name", "pass.type.name", "pass.body_part.name", "pass.technique.name",
                    "pass.technique.id", "pass.assisted_shot_id", "pass.deflected", 
                    "carry.end_location.x", "carry.end_location.y", "ball_receipt.outcome.id",
                    "ball_receipt.outcome.name", "freeze_frame", "visible_area", "group_id")

xgboost_df <- events_rel %>% select(-all_of(vars_to_remove))

# Make any logical column into a 1 or 0
xgboost_df <- xgboost_df %>% mutate_if(is.logical, as.integer)

# Fill all NA with 0
xgboost_df[is.na(xgboost_df)] <- 0

# Fit xgboost model
xgb_model <- xgboost(data = as.matrix(xgboost_df %>% select(-pass.completed, -id)), 
                      label = xgboost_df$pass.completed, 
                      nrounds = 50, 
                      objective = "binary:logistic", 
                      eval_metric = "logloss", 
                      verbose = 1)

# Get predicted probabilities
xgboost_df$passcomp_prob <- predict(xgb_model, as.matrix(xgboost_df %>% select(-pass.completed,-id)))

# Add predicted probabilities to events_rel
events_rel <- events_rel %>% left_join(xgboost_df %>% select(id, passcomp_prob), by = "id")


```


# Get information about the previous pass
```{r}
# if the previous rows pass.end_location.x and pass.end_location.y are the same as the current rows recipient.location.x and recipient.location.y, then create a new variable for previous pass length
events_rel <- events_rel %>% 
  group_by(match_id) %>%
  mutate(prev_pass_length = ifelse(lag(pass.end_location.x) == recipient.location.x & 
                                     lag(pass.end_location.y) == recipient.location.y, 
                                   lag(pass.length), NA),
         prev_pass_angle = ifelse(lag(pass.end_location.x) == recipient.location.x & 
                                     lag(pass.end_location.y) == recipient.location.y,
                                   lag(pass.angle), NA),
         prev_pass_height = ifelse(lag(pass.end_location.x) == recipient.location.x & 
                                     lag(pass.end_location.y) == recipient.location.y,
                                   lag(pass.height.name), NA),
         prev_pass_type = ifelse(lag(pass.end_location.x) == recipient.location.x & 
                                     lag(pass.end_location.y) == recipient.location.y,
                                   lag(pass.type.name), NA),
         prev_pass_body_part = ifelse(lag(pass.end_location.x) == recipient.location.x & 
                                     lag(pass.end_location.y) == recipient.location.y,
                                   lag(pass.body_part.name), NA), 
         prev_pass_switch = ifelse(lag(pass.end_location.x) == recipient.location.x & 
                                     lag(pass.end_location.y) == recipient.location.y,
                                   lag(pass.switch), NA),
         prev_pass_aerial_won = ifelse(lag(pass.end_location.x) == recipient.location.x & 
                                     lag(pass.end_location.y) == recipient.location.y,
                                   lag(pass.aerial_won), NA),
         prev_pass_cross = ifelse(lag(pass.end_location.x) == recipient.location.x & 
                                     lag(pass.end_location.y) == recipient.location.y,
                                   lag(pass.cross), NA),
         prev_pass_cut_back = ifelse(lag(pass.end_location.x) == recipient.location.x & 
                                     lag(pass.end_location.y) == recipient.location.y,
                                   lag(pass.cut_back), NA),
         prev_pass_through_ball = ifelse(lag(pass.end_location.x) == recipient.location.x & 
                                     lag(pass.end_location.y) == recipient.location.y,
                                   lag(pass.through_ball), NA),
         prev_pass_under_pressure = ifelse(lag(pass.end_location.x) == recipient.location.x & 
                                     lag(pass.end_location.y) == recipient.location.y,
                                   lag(under_pressure), NA),
         prev_pass_totalTimeInPoss = ifelse(lag(pass.end_location.x) == recipient.location.x & 
                                     lag(pass.end_location.y) == recipient.location.y,
                                   lag(totalTimeInPoss), NA),
         prev_pass_carried = ifelse(lag(pass.end_location.x) == recipient.location.x & 
                                     lag(pass.end_location.y) == recipient.location.y,
                                   lag(carried), NA),
         prev_passcomp_prob = ifelse(lag(pass.end_location.x) == recipient.location.x & 
                                     lag(pass.end_location.y) == recipient.location.y,
                                   lag(passcomp_prob), NA)) %>%
  ungroup()

# Keep only plays with a previous pass
events_rel <- events_rel %>% filter(!is.na(prev_pass_length))


```


# Extract coordinates from visible_area getting xmin, xmax, ymin and ymax from the list
```{r}
events_rel <- events_rel %>%
  mutate(
    xmin = map_dbl(visible_area, ~ {
      coords <- .x  # assuming .x is a vector of coordinates
      if (length(coords) == 0) {
        NA_real_  # return NA if the vector is empty
      } else {
        x_values <- coords[seq(1, length(coords), by = 2)]
        if (length(x_values) == 0) {
          NA_real_  # return NA if no x-values are extracted
        } else {
          min(x_values)
        }
      }
    }),
    xmax = map_dbl(visible_area, ~ {
      coords <- .x  # assuming .x is a vector of coordinates
      if (length(coords) == 0) {
        NA_real_  # return NA if the vector is empty
      } else {
        x_values <- coords[seq(1, length(coords), by = 2)]
        if (length(x_values) == 0) {
          NA_real_  # return NA if no x-values are extracted
        } else {
          max(x_values)
        }
      }
    }),
    ymin = map_dbl(visible_area, ~ {
      coords <- .x  # assuming .x is a vector of coordinates
      if (length(coords) == 0) {
        NA_real_  # return NA if the vector is empty
      } else {
        y_values <- coords[seq(2, length(coords), by = 2)]
        if (length(y_values) == 0) {
          NA_real_  # return NA if no y-values are extracted
        } else {
          min(y_values)
        }
      }
    }),
    ymax = map_dbl(visible_area, ~ {
      coords <- .x  # assuming .x is a vector of coordinates
      if (length(coords) == 0) {
        NA_real_  # return NA if the vector is empty
      } else {
        y_values <- coords[seq(2, length(coords), by = 2)]
        if (length(y_values) == 0) {
          NA_real_  # return NA if no y-values are extracted
        } else {
          max(y_values)
        }
      }
    })
  )

```


# Standardize so offenses are always attacking from left to right
```{r}
events_ffs <- events_rel %>%
  mutate(ff_id = row_number()) %>%
  ungroup() %>%
  unnest(freeze_frame) %>%
  mutate(ff_location.x = (map(location, 1)), ff_location.y = (map(location, 2))) %>%
  select(-location) %>%
  mutate(
    ff_location.x = as.numeric(ifelse(
      ff_location.x == "NULL", NA, ff_location.x
    )),
    ff_location.y = as.numeric(ifelse(ff_location.y == "NULL", NA, ff_location.y))
  ) #%>%
  # group_by(match_id, period) %>%
  # mutate(AttackingRightToLeft = ifelse(
  #   sum(ff_location.x > 60 & minute == min(minute) & second == min(second[minute == min(minute)])) >
  #     sum(ff_location.x < 60 & minute == min(minute) & second == min(second[minute == min(minute)])),
  #   1,
  #   0
  # )) %>%
  # mutate_at(vars(contains(".x")), funs(ifelse(AttackingRightToLeft == 1 & period == 1, 120 - ., .))) %>%
  # mutate_at(vars(contains(".y")), funs(ifelse(AttackingRightToLeft == 1 & period == 1, 80 - ., .))) %>%
  # mutate_at(vars(contains(".x")), funs(ifelse(AttackingRightToLeft == 0 & period == 2, 120 - ., .))) %>%
  # mutate_at(vars(contains(".y")), funs(ifelse(AttackingRightToLeft == 0 & period == 2, 80 - ., .))) %>%
  # ungroup() 
```

# Calculate the distance of the closest teammate == FALSE where actor == TRUE
```{r}
# Calculate distance from ff_location.x and ff_location.y to location.x and location.y
def_dist_ang <- events_ffs %>%
  mutate(
    distance = sqrt((ff_location.x - location.x)^2 + (ff_location.y - location.y)^2),
    angle = atan2(location.y - ff_location.y, location.x - ff_location.x) * 180 / pi
  ) %>%
  group_by(id) %>%
  summarise(
    closest_defender_distance = min(distance[teammate == FALSE]),
    closest_defender_angle = angle[distance == min(distance[teammate == FALSE])]
  )  %>% 
  filter(closest_defender_distance != Inf) %>%
  ungroup() 

# join the distance and angle to the events_rel
events_rel <- left_join(events_rel, def_dist_ang, by = "id")


```


# Calculate Voronoi Diagrams and Delaunay Triangulation
```{r}
# Perform tessellation for each unique ID
tesselations <- events_ffs %>%
  group_by(id) %>%
  summarise(
    tess_summary = list(tryCatch({
      tesselation <- deldir(ff_location.x, ff_location.y, 
                            rw = c(min(xmin), min(xmax), min(ymin), min(ymax)))
      tesselation$summary
    }, error = function(e) {data.frame()}))
  ) %>%
  ungroup() 

events_vor <- left_join(events_rel, tesselations, by = "id")

events_vor <- events_vor %>%
  mutate(tess_summary = map(tess_summary, ~ as.data.frame(.x))) %>%
  unnest(tess_summary) %>%
  mutate(distance = sqrt((location.x - x)^2 + (location.y - y)^2)) %>%
  filter(distance < 5) %>%
  group_by(id) %>%
  slice_min(order_by = distance, with_ties = FALSE) %>%
  ungroup()

```

# Make column for opposition team
```{r}
# possession_team.id and possession_team.name for the match_id that isn't the one in the row
events_vor <- events_vor %>% distinct(match_id, possession_team.id, possession_team.name) %>%
  rename(opposition_team.id = possession_team.id, opposition_team.name = possession_team.name) %>%
  right_join(events_vor, by = "match_id") %>% filter(possession_team.id != opposition_team.id) 

```

# Select columns to transport
```{r}
big_cols <- c(
  "id",  "totalTimeInPoss",
   "possession_team.name", "opposition_team.name",
  "player.id", "player.name", "position.id", "position.name", "play_pattern.name",
  "location.x", "location.y", "pass.length", "pass.angle", "pass.end_location.x", "pass.end_location.y",
   "pass.height.name",  "pass.body_part.name", "pass.switch",
  "pass.aerial_won", "pass.cross", "pass.cut_back", "pass.through_ball",
  "under_pressure", "passcomp_prob", "recipient.location.x", "recipient.location.y",
  "prev_pass_length", "prev_pass_angle", "prev_pass_height", "prev_pass_height", 
  "prev_pass_body_part", "prev_pass_switch", "prev_pass_aerial_won", "prev_pass_cross", 
  "prev_pass_cut_back", "prev_pass_through_ball", "prev_pass_under_pressure", 
  "prev_pass_totalTimeInPoss", "prev_passcomp_prob",
  "closest_defender_distance", "closest_defender_angle", 
  "n.tri", "del.area", "del.wts", "n.tside", "nbpt", "dir.area", "dir.wts"
)

final_df <- events_vor %>% select(all_of(big_cols))

```

# Count the number of missing values in each column
```{r}
missing_values <- final_df %>% summarise_all(funs(sum(is.na(.)))) %>% gather() %>% arrange(desc(value))

```

# Fill in prev_pass_body_part and pass.body_part.name with "Right Foot"
```{r}
final_df <- final_df %>% mutate(
  prev_pass_body_part = ifelse(is.na(prev_pass_body_part), "Right Foot", prev_pass_body_part),
  pass.body_part.name = ifelse(is.na(pass.body_part.name), "Right Foot", pass.body_part.name)
)

```

# Remove rows with no closest_defender_distance
```{r}
final_df <- final_df %>% filter(!is.na(closest_defender_distance))

```

# Remove missing del.wts
```{r}
final_df <- final_df %>% filter(!is.na(del.wts))

```

# If column is logical make NA FALSE
```{r}
final_df <- final_df %>% mutate_if(is.logical, ~ replace(., is.na(.), FALSE)) 

```

# Column renaming for organization
```{r}
final_df <- final_df %>%
  rename(
    player_id = player.id,
    player_name = player.name,
    position_id = position.id,
    position_name = position.name,
    possession_team = possession_team.name,
    opposition_team = opposition_team.name,
    pattern_name = play_pattern.name,
    
    location_x = location.x,
    location_y = location.y,
    pass_length = pass.length,
    pass_angle = pass.angle,
    pass_end_x = pass.end_location.x,
    pass_end_y = pass.end_location.y,
    
    pass_height = pass.height.name,
    pass_body_part = pass.body_part.name,
    pass_is_switch = pass.switch,
    pass_is_aerial_won = pass.aerial_won,
    pass_is_cross = pass.cross,
    pass_is_cut_back = pass.cut_back,
    pass_is_through_ball = pass.through_ball,
    
    under_pressure = under_pressure,
    pass_completion_probability = passcomp_prob,
    
    recipient_x = recipient.location.x,
    recipient_y = recipient.location.y,
    
    prev_pass_length = prev_pass_length,
    prev_pass_angle = prev_pass_angle,
    prev_pass_height = prev_pass_height,
    prev_pass_body_part = prev_pass_body_part,
    prev_pass_is_switch = prev_pass_switch,
    prev_pass_is_aerial_won = prev_pass_aerial_won,
    prev_pass_is_cross = prev_pass_cross,
    prev_pass_is_cut_back = prev_pass_cut_back,
    prev_pass_is_through_ball = prev_pass_through_ball,
    prev_pass_under_pressure = prev_pass_under_pressure,
    prev_pass_time_in_possession = prev_pass_totalTimeInPoss,
    prev_pass_completion_probability = prev_passcomp_prob,
    
    closest_defender_dist = closest_defender_distance,
    closest_defender_ang = closest_defender_angle,
    
    num_triangles = n.tri,
    delaunay_area = del.area,
    delaunay_weights = del.wts,
    num_side = n.tside,
    num_points= nbpt,
    dirichlet_area = dir.area,
    dirichlet_weights = dir.wts
  )

```


# Write csv
```{r}
write_csv(final_df, "cleaned_WorldCup_passes.csv")

```




