---
title: "Modeling"
author: "Shane Hauck"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load packages
```{r}
library(tidyverse)
library(glmnet)
library(lme4)
library(broom.mixed)

```

# Load data
```{r}
#full_df <- read_csv("cleaned_WorldCup_passes.csv")
epl_df <- read_csv("EPL_1516_passes.csv") %>% mutate(league = "EPL")
la_liga_df <- read_csv("LaLiga_1516_passes.csv") %>% mutate(league = "La Liga")
serie_a_df <- read_csv("SerieA_1516_passes.csv") %>% mutate(league = "Serie A")
bundesliga_df <- read_csv("Bundesliga_1516_passes.csv") %>% mutate(league = "Bundesliga")
ligue_1_df <- read_csv("Ligue1_1516_passes.csv") %>% mutate(league = "Ligue 1")

full_df <- rbind(epl_df, la_liga_df, serie_a_df, bundesliga_df, ligue_1_df)


```

# Group positions
```{r}
full_df <- full_df %>%
  mutate(
    position = case_when(
      grepl("Goalkeeper", position_name) ~ "Goalkeeper",
      grepl("Center Back", position_name) ~ "Center Back",
      grepl("Left Back", position_name) ~ "Full Back",
      grepl("Right Back", position_name) ~ "Full Back",
      grepl("Wing Back", position_name) ~ "Wing Back",
      grepl("Defensive Midfield", position_name) ~ "Defensive Midfield",
      grepl("Center Midfield", position_name) ~ "Center Midfield",
      grepl("Attacking Midfield", position_name) ~ "Attacking Midfield",
      grepl("Left Wing", position_name) ~ "Winger",
      grepl("Right Wing", position_name) ~ "Winger",
      grepl("Right Midfield", position_name) ~ "Winger",
      grepl("Left Midfield", position_name) ~ "Winger",
      grepl("Forward", position_name) ~ "Forward"
    )
  ) %>%
  select(-position_name)
```


# Filter data to be x between 20 and 100
```{r}
full_df <- full_df %>%
  filter(location_x >= 20 & location_x <= 100)

```

# Filter to keep passes with body part including Foot
```{r}
full_df <- full_df %>%
  filter(str_detect(pass_body_part, "Foot"))

```


# Square root timeInPoss
```{r}
full_df <- full_df %>%
  mutate(totalTimeInPoss_sqrt = sqrt(totalTimeInPoss))

```

# Standardize numeric columns
```{r}
num_cols <- full_df %>% select_if(is.numeric) %>% select(-totalTimeInPoss_sqrt) %>% colnames()

# full_df <- full_df %>%
#   mutate_at(vars(num_cols), (. - min(.) ) / (max(.) - min(.)))

```


# Columns for LASSO model
```{r}
rmv_cols <- c(
  "possession_team", "opposition_team", "player_id", "player_name", "position_id", "totalTimeInPoss", "match_id"
)

lasso_df <- full_df %>%
  select(-all_of(rmv_cols)) %>% 
  select(id, totalTimeInPoss_sqrt, everything())

```

# Fit LASSO model to predict totalTimeInPoss_sqrt to extract important features
```{r}
set.seed(12)
lasso_model <- cv.glmnet(
  x = as.matrix(lasso_df %>% select(-id, -totalTimeInPoss_sqrt) %>% select_if(is.numeric)),
  y = lasso_df$totalTimeInPoss_sqrt,
  alpha = 1,
  family = "gaussian",
  nfolds = 10
)

plot(lasso_model)

# Extract coefficients at lambda.min
coefficients <- coef(lasso_model, s = "lambda.min")

# Coerce to a regular matrix and check non-zero entries
coefficients_matrix <- as.matrix(coefficients)
non_zero_indices <- which(coefficients_matrix[,1] != 0)

# Extract names of non-zero coefficients
non_zero_coefficients <- rownames(coefficients_matrix)[non_zero_indices]

# Remove intercept
non_zero_coefficients <- non_zero_coefficients[-1]

```

# Fit model with training testing on RMSE with test
```{r}
set.seed(12)
train_idx <- sample(1:nrow(lasso_df), 0.8*nrow(lasso_df))
train_data <- lasso_df[train_idx,]
test_data <- lasso_df[-train_idx,]

x_train <- as.matrix(train_data %>% select(-id, -totalTimeInPoss_sqrt))
y_train <- train_data$totalTimeInPoss_sqrt
x_test <- as.matrix(test_data %>% select(-id, -totalTimeInPoss_sqrt))
y_test <- test_data$totalTimeInPoss_sqrt

# Fit LASSO model using cross-validation
lasso_model <- cv.glmnet(
  x = x_train,
  y = y_train,
  alpha = 1,
  family = "gaussian",
  nfolds = 10
)

# Plot the cross-validated LASSO model
plot(lasso_model)

# Predict on the test dataset using the lambda that gives minimum mean cross-validated error
predicted_values <- predict(lasso_model, s = "lambda.min", newx = x_test)

# Calculate RMSE
rmse <- sqrt(mean((y_test - predicted_values)^2))
print(paste("RMSE:", rmse))

# Plotting predicted vs observed values
obs_vs_pred <- data.frame(Observed = y_test, Predicted = as.vector(predicted_values))
ggplot(obs_vs_pred, aes(x = Predicted, y = Observed)) +
  geom_point(alpha = 0.25) +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  labs(title = "Observed vs Predicted values",
       x = "Predicted",
       y = "Observed")

```

# Fit Multilevel model with player effects modeling time in possession until pass
```{r}
# Create formula with non-zero coefficients
formula <- as.formula(paste("totalTimeInPoss_sqrt ~", 
                            paste(non_zero_coefficients, collapse = " + "),
                            " + (1 | player_id) + (1 | position) + (1 | opposition_team) + (1 | possession_team) +
                            (1 | league)"))

# Fit multilevel model
time_mod <- lmer(
  formula = formula,
  data = full_df#,
 # family = Gamma(link = "log")
)
summary(time_mod)

# Extract fixed effects
fixed_effects_time <- tidy(time_mod, effects = "fixed")

# Extract random effects
random_effects_time <- tidy(time_mod, effects = "ran_vals")
```

# Plot Resiudals

```{r}
# Extract residuals
residuals <- residuals(time_mod)

# Plot residuals
ggplot(data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(bins = 30) +
  labs(title = "Residuals of the model",
       x = "Residuals",
       y = "Frequency")

```

# Join random effects with player names
```{r}
player_names <- full_df %>%
  select(player_id, player_name, position, possession_team) %>%
  group_by(player_id) %>%
  mutate(count = n()) %>%
  ungroup() %>% 
  distinct(player_id, .keep_all = T)

random_effects_time <- random_effects_time %>%
  #separate(level, into = c("player_id", "position"), sep = ":") %>%
  left_join(player_names %>% mutate(player_id = as.character(player_id)),
            by = c("level" = "player_id"))

```


# Calculate Intraclass Correlation Coefficient
```{r}
VarCorr(time_mod) %>%
  as_tibble() %>%
  mutate(ICC = vcov / sum(vcov)) %>%
  select(grp, ICC)
```


# Bootstrap to get confidence intervals
```{r}
# Empty dataframe to store bootstrap results
bootstrap_results_time <- data.frame()

bootstrap_icc_time <- data.frame()

# Number of bootstrap samples
n_samples <- 100

# Bootstrap loop
set.seed(12)
for(i in 1:n_samples){
  # Sample with replacement
  sample_idx <- sample(1:nrow(full_df), replace = T)
  sample_df <- full_df[sample_idx,]
  
  # Fit model
  time_mod_sample <- lmer(
    formula = formula,
    data = sample_df#,
    #family = Gamma(link = "log")
  )
  
  # Extract fixed effects
  fixed_effects_sample <- tidy(time_mod_sample, effects = "fixed")
  
  # Extract random effects
  random_effects_sample <- tidy(time_mod_sample, effects = "ran_vals")
  
  # Store results
  bootstrap_results_time <- rbind(bootstrap_results_time, 
                                  rbind(fixed_effects_sample %>% mutate(level = NA, group = NA), 
                                        random_effects_sample %>% mutate(statistic = NA)) %>% mutate(boot = i))
  
  # Calculate ICC
  icc_sample <- VarCorr(time_mod_sample) %>%
    as_tibble() %>%
    mutate(ICC = vcov / sum(vcov)) %>%
    select(grp, ICC) %>%
    mutate(boot = i)
  
  bootstrap_icc_time <- rbind(bootstrap_icc_time, icc_sample)
}

# Join with player names
bootstrap_results_time <- bootstrap_results_time %>%
  left_join(player_names %>% mutate(player_id = as.character(player_id)),
            by = c("level" = "player_id"))

# Write to csv
write_csv(bootstrap_results_time, "bootstrap_results_time.csv")

write_csv(bootstrap_icc_time, "bootstrap_icc_time.csv")


```




# Filter for opposition_team
```{r}
opposition_df <- random_effects_time %>%
  filter(str_detect(group, "opposition_team"))
```

# Filter for possession_team
```{r}
possession_df <- random_effects_time %>%
  filter(str_detect(group, "possession_team"))
```

# Filter for league
```{r}
league_df <- random_effects_time %>%
  filter(str_detect(group, "league"))

```

# Filter for position
```{r}
position_df <- random_effects_time %>%
  filter(group == "position")

```


# Players whose estimate is closest to 0
```{r}
random_effects_time %>%
  filter(count > 300) %>%
  arrange(abs(estimate)) %>%
  head(100) %>% view()

```


# Create proportion of total passes for both teams that a player plays in a game
```{r}
full_df <- full_df %>% group_by(match_id) %>% 
  mutate(match_npasses = n()) %>%
  group_by(match_id, player_id) %>%
  mutate(player_npasses = n()) %>% 
  ungroup() %>%
  mutate(scaled_player_npasses = (player_npasses - min(player_npasses))/(max(player_npasses) - min(player_npasses))) %>%
  mutate(prop_passes = player_npasses/match_npasses) %>%
  mutate(scaled_prop_passes = sqrt(prop_passes * scaled_player_npasses)) %>%
  ungroup() 
  
```

# Multilevel model for scaled proportion of passes
```{r}
# Fit multilevel model
prop_mod <- lmer(prop_passes ~ (1 | player_id) + (1 | position) + 
                   (1 | opposition_team) + (1 | possession_team), 
                 data = full_df)

```

# Extract random effects
```{r}
random_effects_prop <- tidy(prop_mod, effects = "ran_vals") %>%
  left_join(player_names %>% mutate(player_id = as.character(player_id)),
            by = c("level" = "player_id"))

```

# Calculate ICC
```{r}
VarCorr(prop_mod) %>%
  as_tibble() %>%
  mutate(ICC = vcov / sum(vcov)) %>%
  select(grp, ICC)
```

# Bootstrap to get confidence intervals
```{r}
# Empty dataframe to store bootstrap results
bootstrap_results_prop <- data.frame()

bootstrap_icc_prop <- data.frame()

# Number of bootstrap samples
n_samples <- 100

# Bootstrap loop
set.seed(12)
for(i in 1:n_samples){
  # Sample with replacement
  sample_idx <- sample(1:nrow(full_df), replace = T)
  sample_df <- full_df[sample_idx,]
  
  # Fit model
  prop_mod_sample <- lmer(
    formula = prop_passes ~ (1 | player_id) + (1 | position) + 
      (1 | opposition_team) + (1 | possession_team),
    data = sample_df
  )
  
  # Extract fixed effects
  fixed_effects_sample <- tidy(prop_mod_sample, effects = "fixed")
  
  # Extract random effects
  random_effects_sample <- tidy(prop_mod_sample, effects = "ran_vals")
  
  # Store results
  bootstrap_results_prop <- rbind(bootstrap_results_prop, 
                                  rbind(fixed_effects_sample %>% mutate(level = NA, group = NA), 
                                        random_effects_sample %>% mutate(statistic = NA)) %>% mutate(boot = i))
  
  # Calculate ICC
  icc_sample <- VarCorr(prop_mod_sample) %>%
    as_tibble() %>%
    mutate(ICC = vcov / sum(vcov)) %>%
    select(grp, ICC) %>%
    mutate(boot = i)
  
  bootstrap_icc_prop <- rbind(bootstrap_icc_prop, icc_sample)
}

# Join with player names
bootstrap_results_prop <- bootstrap_results_prop %>%
  left_join(player_names %>% mutate(player_id = as.character(player_id)),
            by = c("level" = "player_id"))

# Write to csv
write_csv(bootstrap_results_prop, "bootstrap_results_prop.csv")

write_csv(bootstrap_icc_prop, "bootstrap_icc_prop.csv")



```


```{r}
player_re_t <- random_effects_time %>%
  filter(group == "player_id") %>%
  rename(
    estimate_time = estimate,
    std.error_time = std.error
  ) 

player_re_p <- random_effects_prop %>%
  filter(group == "player_id") %>%
  rename(
    estimate_prop = estimate,
    std.error_prop = std.error
  ) %>% select(level, estimate_prop, std.error_prop)

left_join(player_re_t, player_re_p)  %>%
  select(level, player_name, position, possession_team, estimate_time, std.error_time, estimate_prop, std.error_prop, count) %>%
  mutate(scaled_estimate_prop = (estimate_prop - min(estimate_prop))/(max(estimate_prop) - min(estimate_prop)))  %>%
  mutate(estimate_full = estimate_time * exp(3 * scaled_estimate_prop)) %>% 
  arrange(abs(estimate_full)) %>%
  view()

```


# Ridge plots with bootstrapped results time
```{r}
library(ggridges)

bootstrap_results_time <- read_csv("bootstrap_results_time.csv")

# Top 10 players with the smallest estimate
fast_players_10 <- bootstrap_results_time %>%
  filter(!is.na(level)) %>%
  filter(position != "Goalkeeper") %>%
  group_by(level, player_name, position, possession_team) %>% 
  summarise(median_estimate = median(estimate),
            lower = quantile(estimate, 0.05),
            upper = quantile(estimate, 0.95)
            
              ) %>%
  arrange(median_estimate) %>%
  head(10) 

slow_players_10 <- bootstrap_results_time %>%
  filter(!is.na(level)) %>%
  filter(position != "Goalkeeper") %>%
  group_by(level, player_name, position, possession_team) %>% 
  summarise(median_estimate = median(estimate),
            lower = quantile(estimate, 0.05),
            upper = quantile(estimate, 0.95)
            
              ) %>%
  arrange(desc(median_estimate)) %>%
  head(10)

# list of top 5 and bottom 5
top_bottom_players <- rbind(fast_players_10 %>% head(5), slow_players_10 %>% head(5)) %>% arrange(median_estimate)

# Ridge plot
bootstrap_results_time %>%
  filter(level %in% top_bottom_players$level) %>%
  group_by(level) %>% 
  mutate(median_estimate = median(estimate)) %>%
  ungroup() %>%
  ggplot(aes(x = estimate, y = reorder(player_name, desc(median_estimate)), fill = median_estimate)) +
  geom_density_ridges(alpha = 0.7, scale = .97,
                      quantile_lines = TRUE, quantiles = 2, 
                      rel_min_height = 0.01) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_bw() +
  theme(legend.position = "none") +
  labs(title = "Bootstrap Distribution of Player Effects on Time in Possession",
       subtitle = "Top 5 Fastest and Slowest Passers",
       x = "Rhythm Estimate",
       y = "Player") +
  scale_fill_viridis_c(option = "turbo", direction = -1) +
  theme(axis.text.y = element_text(size = 10)) +
  theme(axis.text.x = element_text(size = 10)) +
  theme(plot.title = element_text(size = 16, face = "bold")) +
  theme(plot.subtitle = element_text(size = 10, face = "bold")) +
  theme(axis.title = element_text(size = 10, face = "bold"))
```


# Ridge plots with bootstrapped results prop
```{r}
bootstrap_results_prop <- read_csv("bootstrap_results_prop.csv")

# Top 10 players with the largest estimate
control_players_10 <- bootstrap_results_prop %>%
  filter(!is.na(level)) %>%
  filter(position != "Goalkeeper") %>%
  group_by(level, player_name, position, possession_team) %>% 
  summarise(median_estimate = median(estimate),
            lower = quantile(estimate, 0.05),
            upper = quantile(estimate, 0.95)
            
              ) %>%
  arrange(desc(median_estimate)) %>%
  head(10)


# Ridge plot
bootstrap_results_prop %>%
  filter(level %in% control_players_10$level) %>%
  group_by(level) %>% 
  mutate(median_estimate = median(estimate)) %>%
  ungroup() %>%
  ggplot(aes(x = estimate, y = reorder(player_name, (median_estimate)), fill = median_estimate)) +
  geom_density_ridges(alpha = 0.7, scale = .97,
                      quantile_lines = TRUE, quantiles = 2, 
                      rel_min_height = 0.01) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(title = "Bootstrap Distribution of Player Effects on Proportion of Game Passes",
       subtitle = "Top 10 Players with Most Control Over Game Passing Flow",
       x = "Control Estimate",
       y = "Player") +
  scale_fill_viridis_c(option = "magma", direction = -1) +
  theme(axis.text.y = element_text(size = 10)) +
  theme(axis.text.x = element_text(size = 10)) +
  theme(plot.title = element_text(size = 16, face = "bold")) +
  theme(plot.subtitle = element_text(size = 10, face = "bold")) +
  theme(axis.title = element_text(size = 10, face = "bold"))

```


# Combining Rhythm and Control
```{r}
player_re_t <- bootstrap_results_time %>%
  filter(group == "player_id") %>%
  rename(
    estimate_time = estimate,
    std.error_time = std.error
  ) 

player_re_p <- bootstrap_results_prop %>%
  filter(group == "player_id") %>%
  rename(
    estimate_prop = estimate,
    std.error_prop = std.error
  ) %>% select(level, estimate_prop, std.error_prop, boot)

RhythmControlScores <- left_join(player_re_t, player_re_p, by = c("level", "boot"))  %>%
  filter(position != "Goalkeeper") %>%
  filter(player_name != "Mirko Valdifiori") %>%
  select(level, player_name, position, possession_team, estimate_time, 
         std.error_time, estimate_prop, std.error_prop, count) %>%
  ungroup() %>%
  mutate(scaled_estimate_prop = (estimate_prop - min(estimate_prop))/(max(estimate_prop) - min(estimate_prop)))  %>%
  mutate(estimate_full = estimate_time * exp(3 * scaled_estimate_prop)) %>% 
  group_by(player_name, position, possession_team) %>%
  summarise(median_estimate_full = median(estimate_full),
            lower = quantile(estimate_full, 0.05),
            upper = quantile(estimate_full, 0.95)) %>%
  ungroup()

write_csv(RhythmControlScores, "RhythmControlScores.csv")
  
# Top 10 players with the largest estimate
slow_control_players_10 <- RhythmControlScores %>%
  arrange(desc(median_estimate_full)) %>%
  head(10)

fast_control_players_10 <- RhythmControlScores %>%
  arrange(median_estimate_full) %>%
  head(10)

top_bottom_players <- bind_rows(slow_control_players_10 %>% head(5), fast_control_players_10 %>% head(5))

# Ridge plot
left_join(player_re_t, player_re_p, by = c("level", "boot"))  %>%
  filter(position != "Goalkeeper") %>%
  filter(player_name != "Mirko Valdifiori") %>%
  select(level, player_name, position, possession_team, estimate_time, 
         std.error_time, estimate_prop, std.error_prop, count) %>%
  mutate(scaled_estimate_prop = (estimate_prop - min(estimate_prop))/(max(estimate_prop) - min(estimate_prop)))  %>%
  mutate(estimate_full = estimate_time * exp(3 * scaled_estimate_prop)) %>%
  filter(player_name %in% top_bottom_players$player_name) %>%
  group_by(player_name, position, possession_team) %>%
  mutate(median_estimate_full = median(estimate_full)) %>%
  ggplot(aes(x = estimate_full, y = reorder(player_name, desc(median_estimate_full)), fill = median_estimate_full)) +
  geom_density_ridges(alpha = 0.7, scale = .97,
                      quantile_lines = TRUE, quantiles = 2, 
                      rel_min_height = 0.01) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(title = "Bootstrap Distribution of Player Effects Rhythm Control",
       subtitle = "Top 5 Fastest and Slowest Passers with Most Control Over Game Passing Flow",
       x = "Rhythm Control Score",
       y = "Player") +
  scale_fill_viridis_c(option = "turbo", direction = -1) +
  theme(axis.text.y = element_text(size = 10)) +
  theme(axis.text.x = element_text(size = 10)) +
  theme(plot.title = element_text(size = 16, face = "bold")) +
  theme(plot.subtitle = element_text(size = 10, face = "bold")) +
  theme(axis.title = element_text(size = 10, face = "bold"))


```


```{r}
df <- player_re_t %>%
  left_join(player_re_p, by = c("level", "boot")) %>%
  filter(position != "Goalkeeper", player_name != "Mirko Valdifiori") %>%
  select(level, player_name, position, possession_team, estimate_time, 
         std.error_time, estimate_prop, std.error_prop, count) %>%
  ungroup() %>%
  mutate(scaled_estimate_prop = (estimate_prop - min(estimate_prop))/(max(estimate_prop) - min(estimate_prop))) %>%
  mutate(estimate_full = estimate_time * exp(3 * scaled_estimate_prop)) %>% 
  group_by(player_name, position, possession_team) %>%
  summarise(median_estimate_full = median(estimate_full),
            median_estimate_time = median(estimate_time),
            median_estimate_prop = median(estimate_prop),
            lower = quantile(estimate_full, 0.05),
            upper = quantile(estimate_full, 0.95)) %>%
  filter(median_estimate_prop > 0.025)

# Create the ggplot
p <- ggplot(df, aes(x = median_estimate_prop, y = median_estimate_time, color = median_estimate_full)) +
  geom_point(alpha = 0.5, size = 3) +
  theme_bw() +
  scale_color_gradientn(colors = c("orange", "purple", "blue")) +
  ggtitle("Player Performance Estimates")

# Convert to plotly and add custom hover info
plotly_fig <- plot_ly(data = df, x = ~median_estimate_prop, y = ~median_estimate_time, 
                      type = 'scatter', mode = 'markers',
                      marker = list(size = 10, color = ~median_estimate_full, 
                                    colorscale = list( c(-1.36, "red"),   # Low values in red
                                                       c(1.16, "blue")   # High values in blue
                                                       )),
                      text = ~paste('Player Name: ', player_name, 
                                    '<br>Rhythm Control Score: ', round(median_estimate_full,2),
                                    '<br>Rhythm Effect: ', round(median_estimate_time,2),
                                    '<br>Control Effect: ', round(median_estimate_prop,2)),
                      hoverinfo = 'text') %>%
  layout(title = 'Player Rhythm and Control',
         xaxis = list(title = 'Control Effect', size = 14),
         yaxis = list(title = 'Rhythm Effect', size = 14),
         legend = list(title = 'Rhythm Control Score'),
         hoverlabel = list(bgcolor = 'white')) 


plotly_fig




```






<!-- # Fit Bayesian Multilevel model with player effects modeling time in possession until pass -->
<!-- ```{r} -->
<!-- library(rstanarm) -->
<!-- ``` -->

<!-- # Make player_id a factor -->
<!-- ```{r} -->
<!-- full_df$player_id <- as.factor(full_df$player_id) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- # Create formula with non-zero coefficients -->
<!-- formula <- as.formula(paste("totalTimeInPoss_sqrt ~",  -->
<!--                             paste(non_zero_coefficients, collapse = " + "), -->
<!--                             " + (1 | player_id) + (1 | position) + (1 | opposition_team) + (1 | possession_team) + -->
<!--                             (1 | league)")) -->

<!-- # Fit multilevel model -->
<!-- time_mod_bayes <- stan_glmer( -->
<!--   formula = formula, -->
<!--   data = full_df, -->
<!--   # Gamma distribution -->
<!--   family = Gamma(link = "log"), -->
<!--   seed = 12, -->
<!--   iter = 2000, -->
<!--   chains = 4, -->
<!--   cores = 4 -->
<!-- ) -->


<!-- ``` -->

<!-- # Fit Bayesian Multilevel model with player effects modeling proportion of passes -->
<!-- ```{r} -->
<!-- # Fit multilevel model -->
<!-- prop_mod_bayes <- stan_lmer(prop_passes ~ (1 | player_id) + (1 | position) +  -->
<!--                    (1 | opposition_team) + (1 | possession_team),  -->
<!--                  data = full_df, -->
<!--                  seed = 12, -->
<!--                  iter = 500, -->
<!--                  chains = 4, -->
<!--                  cores = 8) -->

<!-- ``` -->


<!-- # Change player_id to be 0-N_players -->
<!-- ```{r} -->
<!-- player_indexs <- full_df %>% -->
<!--   distinct(player_id) %>% -->
<!--   mutate(player_i = row_number()) -->
<!-- position_indexs <- full_df %>% -->
<!--   distinct(position) %>% -->
<!--   mutate(position_i = row_number()) -->
<!-- oppteam_indexs <- full_df %>% -->
<!--   distinct(opposition_team) %>% -->
<!--   mutate(oppteam_i = row_number()) -->
<!-- possteam_indexs <- full_df %>% -->
<!--   distinct(possession_team) %>% -->
<!--   mutate(possteam_i = row_number()) -->
<!-- league_indexs <- full_df %>% -->
<!--   distinct(league) %>% -->
<!--   mutate(league_i = row_number()) -->

<!-- # join with full_df -->
<!-- full_df <- full_df %>% -->
<!--   left_join(player_indexs, by = "player_id") %>% -->
<!--   left_join(position_indexs, by = "position") %>% -->
<!--   left_join(oppteam_indexs, by = "opposition_team") %>% -->
<!--   left_join(possteam_indexs, by = "possession_team") %>% -->
<!--   left_join(league_indexs, by = "league") -->


<!-- ``` -->


<!-- # Possession time model with Stan -->
<!-- ```{r} -->
<!-- library(rstan) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- model_code <- " -->
<!-- data { -->

<!--   int<lower=0> N_passes; // Number of passes (observations) -->
<!--   int N_players; // Number of players -->
<!--   int<lower=0> N_positions; // Number of positions -->
<!--   int<lower=0> N_possteams; // Number of possession teams -->
<!--   int<lower=0> N_oppteams; // Number of opposition teams -->
<!--   int<lower=0> N_leagues; // Number of leagues -->

<!--   // Response Variable -->
<!--   vector[N_passes] totalTimeInPoss_sqrt; // Time in possession -->

<!--   // Fixed Effects -->
<!--   vector[N_passes] location_x; // x-coordinate of the pass -->
<!--   vector[N_passes] pass_length; // Pass length of the pass -->
<!--   vector[N_passes] pass_angle; // Angle of the pass -->
<!--   vector[N_passes] pass_end_x; // x-coordinate of the end of the pass -->
<!--   vector[N_passes] pass_end_y; // y-coordinate of the end of the pass -->
<!--   vector[N_passes] pass_completion_probability; // Completion probability of the pass -->
<!--   vector[N_passes] recipient_x; // Reception x-coordinate of the player -->
<!--   vector[N_passes] recipient_y; // Reception y-coordinate of the player -->
<!--   vector[N_passes] prev_pass_length; // Previous pass length -->
<!--   vector[N_passes] prev_pass_angle; // Previous pass angle -->
<!--   vector[N_passes] prev_pass_time_in_possession; // Previous pass time in possession -->
<!--   vector[N_passes] prev_pass_completion_probability; // Completion probability of the previous pass -->

<!--   int<lower=1, upper=N_players> player_id[N_passes]; // Player ID -->
<!--   int<lower=1, upper=N_positions> position_id[N_passes]; // Position ID -->
<!--   int<lower=1, upper=N_possteams> possteam_id[N_passes]; // Possession team ID -->
<!--   int<lower=1, upper=N_oppteams> oppteam_id[N_passes]; // Opposition team ID -->
<!--   int<lower=1, upper=N_leagues> league_id[N_passes]; // League ID -->

<!-- } -->


<!-- parameters { -->

<!--   vector[N_players] player_effect; // Vector of player intercepts -->
<!--   vector[N_positions] position_effect; // Vector of position intercepts -->
<!--   vector[N_possteams] possteam_effect; // Vector of possession team intercepts -->
<!--   vector[N_oppteams] oppteam_effect; // Vector of opposition team intercepts -->
<!--   vector[N_leagues] league_effect; // Vector of league intercepts -->

<!--   real<lower=0> sigma_players; // Variance of players -->
<!--   real<lower=0> sigma_positions; // Variance of positions -->
<!--   real<lower=0> sigma_possteams; // Variance of possession teams -->
<!--   real<lower=0> sigma_oppteams; // Variance of opposition teams -->
<!--   real<lower=0> sigma_leagues; // Variance of leagues -->

<!--   real mu_players; // Mean of players -->
<!--   real mu_positions; // Mean of positions -->
<!--   real mu_possteams; // Mean of possession teams -->
<!--   real mu_oppteams; // Mean of opposition teams -->
<!--   real mu_leagues; // Mean of leagues -->

<!--   real beta_location_x; // Beta for location_x -->
<!--   real beta_pass_length; // Beta for pass_length -->
<!--   real beta_pass_angle; // Beta for pass_angle -->
<!--   real beta_pass_end_x; // Beta for pass_end_x -->
<!--   real beta_pass_end_y; // Beta for pass_end_y -->
<!--   real beta_pass_completion_probability; // Beta for pass_completion_probability -->
<!--   real beta_recipient_x; // Beta for recipient_x -->
<!--   real beta_recipient_y; // Beta for recipient_y -->
<!--   real beta_prev_pass_length; // Beta for prev_pass_length -->
<!--   real beta_prev_pass_angle; // Beta for prev_pass_angle -->
<!--   real beta_prev_pass_time_in_possession; // Beta for prev_pass_time_in_possession -->
<!--   real beta_prev_pass_completion_probability; // Beta for prev_pass_completion_probability -->

<!--   real<lower=0> sigma_time; // Residual standard deviation -->

<!-- } -->

<!-- model { -->

<!--   // Conditional mean -->
<!--   vector[N_passes] mu_time; -->

<!--   // Linear combination -->
<!--   mu_time = player_effect[player_id] + position_effect[position_id] + possteam_effect[possteam_id] +  -->
<!--       oppteam_effect[oppteam_id] + league_effect[league_id] +  -->
<!--       beta_location_x * location_x + beta_pass_length * pass_length + beta_pass_angle * pass_angle +  -->
<!--       beta_pass_end_x * pass_end_x + beta_pass_end_y * pass_end_y +  -->
<!--       beta_pass_completion_probability * pass_completion_probability + beta_recipient_x * recipient_x +  -->
<!--       beta_recipient_y * recipient_y + beta_prev_pass_length * prev_pass_length +  -->
<!--       beta_prev_pass_angle * prev_pass_angle + beta_prev_pass_time_in_possession * prev_pass_time_in_possession +  -->
<!--       beta_prev_pass_completion_probability * prev_pass_completion_probability; -->

<!--   // Priors -->
<!--   beta_location_x ~ normal(0, 1); -->
<!--   beta_pass_length ~ normal(0, 1); -->
<!--   beta_pass_angle ~ normal(0, 1); -->
<!--   beta_pass_end_x ~ normal(0, 1); -->
<!--   beta_pass_end_y ~ normal(0, 1); -->
<!--   beta_pass_completion_probability ~ normal(0, 1); -->
<!--   beta_recipient_x ~ normal(0, 1); -->
<!--   beta_recipient_y ~ normal(0, 1); -->
<!--   beta_prev_pass_length ~ normal(0, 1); -->
<!--   beta_prev_pass_angle ~ normal(0, 1); -->
<!--   beta_prev_pass_time_in_possession ~ normal(0, 1); -->
<!--   beta_prev_pass_completion_probability ~ normal(0, 1); -->

<!--   // Hyper-priors -->
<!--   mu_players ~ normal(0, 1); -->
<!--   mu_positions ~ normal(0, 1); -->
<!--   mu_possteams ~ normal(0, 1); -->
<!--   mu_oppteams ~ normal(0, 1); -->
<!--   mu_leagues ~ normal(0, 1); -->

<!--   sigma_players ~ cauchy(0, 5); -->
<!--   sigma_positions ~ cauchy(0, 5); -->
<!--   sigma_possteams ~ cauchy(0, 5); -->
<!--   sigma_oppteams ~ cauchy(0, 5); -->
<!--   sigma_leagues ~ cauchy(0, 5); -->

<!--   // Level 2 -->
<!--   player_effect ~ normal(mu_players, sigma_players); -->
<!--   position_effect ~ normal(mu_positions, sigma_positions); -->
<!--   possteam_effect ~ normal(mu_possteams, sigma_possteams); -->
<!--   oppteam_effect ~ normal(mu_oppteams, sigma_oppteams); -->
<!--   league_effect ~ normal(mu_leagues, sigma_leagues); -->

<!--   // Level 1 -->
<!--   totalTimeInPoss_sqrt ~ normal(mu_time, sigma_time); -->

<!-- } -->
<!-- " -->
<!-- ``` -->

<!-- # Create the data list -->
<!-- ```{r} -->
<!-- bayes_time_mod_data <- list( -->
<!--   N_passes = nrow(full_df), -->
<!--   N_players = length(unique(full_df$player_id)), -->
<!--   N_positions = length(unique(full_df$position)), -->
<!--   N_possteams = length(unique(full_df$possession_team)), -->
<!--   N_oppteams = length(unique(full_df$opposition_team)), -->
<!--   N_leagues = length(unique(full_df$league)), -->
<!--   totalTimeInPoss_sqrt = full_df$totalTimeInPoss_sqrt, -->
<!--   location_x = full_df$location_x, -->
<!--   pass_length = full_df$pass_length, -->
<!--   pass_angle = full_df$pass_angle, -->
<!--   pass_end_x = full_df$pass_end_x, -->
<!--   pass_end_y = full_df$pass_end_y, -->
<!--   pass_completion_probability = full_df$pass_completion_probability, -->
<!--   recipient_x = full_df$recipient_x, -->
<!--   recipient_y = full_df$recipient_y, -->
<!--   prev_pass_length = full_df$prev_pass_length, -->
<!--   prev_pass_angle = full_df$prev_pass_angle, -->
<!--   prev_pass_time_in_possession = full_df$prev_pass_time_in_possession, -->
<!--   prev_pass_completion_probability = full_df$prev_pass_completion_probability, -->
<!--   player_id = full_df$player_i, -->
<!--   position_id = full_df$position_i, -->
<!--   possteam_id = full_df$possteam_i, -->
<!--   oppteam_id = full_df$oppteam_i, -->
<!--   league_id = full_df$league_i -->
<!-- )  -->
<!-- ``` -->

<!-- # Fit the model -->
<!-- ```{r} -->
<!-- bayes_time_mod_fit <- stan(model_code = model_code,  -->
<!--                            data = bayes_time_mod_data,  -->
<!--                            iter = 500,  -->
<!--                            chains = 4, -->
<!--                            cores = 4,  -->
<!--                            seed = 28) -->

<!-- ``` -->


